---
title: "Summary report for data challenge"
author: "Eduardo de Freitas Costa"
date: '`r format(Sys.Date(), "%Y-%m-%d")`'
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: yes
    toc_depth: 3
---

```{r}

#header #################################################################################
#'Incidencia8_11.Rmd'

#Title: title
#Project ID: pid
#Client: client
#Author: <Eduardo> <Costa>, UFRGS

#Description: description

#Start date: date
#Last Update: {6:date}

#R version: r.version
#Scriptversion: version

#Dependencies
#<-Downstream
 #-Main.R
#->Upstream


#Input:
#- "Data/Processed/inc_fin.csv"

#Output:
#- "Figures/Fig1.jpg"
#- "Figures/Fig2.jpg"
#- "Output/incidence.csv"

#Peer reviewer(s)

#Please ensure directories are relative. Please comment code sufficiently.

#Script start#############################################################################

```



```{r setup, include=FALSE}


# Define knitr options ########################################################


# global options
knitr::opts_chunk$set(message = FALSE)    # code default outputs
knitr::opts_chunk$set(warning = FALSE)    # code warnings
knitr::opts_chunk$set(echo = TRUE)        # code chunks show
```



```{r, include=FALSE}

# Loading and preparing the data

db <- read.csv(here("Data","Processed" ,"inc_fin.csv"))
db2 <- read.csv(here("Data", "Processed","conf_inc.csv"))

#View(db)
#View(db2)
str(db)
#str(db2)

colnames(db)[1]<-"time_line"

db$time_line<-as.Date(db$time_line)
db2$dataa<-as.Date(db2$dataa)



db3<-cbind.data.frame(
  db$time_line,
  flt01=db$flt01/db$X..Inf.0.*100000,
  f0104=db$f0104/db$X.0.4.*100000,
  f0509=db$f0509/db$X.4.9.*100000,
  f1019=db$f1019/db$X.9.19.*100000,
  f2059=db$f2059/db$X.19.59.*100000,
  f6079=db$f6079/db$X.59.79. *100000,
  fmm80=db$fmm80/db$X.79..Inf.*100000
)

str(db)
#str(db2)

```

# Descriptive
```{r, echo=FALSE}
## Full plots
vac<-c(rep(0,108),rep(1,108))
vac<-ts(vac,start=c(2002,1), end=c(2019, 12), frequency = 12) 


for(i in 2:dim(db)[1]){
  db3$year[1]<-2002
  db3$year[i]<-db3$year[i-1]+(1/12)
}



lab<-c("Before", "After")
col<-c('black','darkgray')

theme_set(theme_minimal())
a<-ggplot(data=db3,aes(x=year,y=flt01,colour=as.factor(vac)))+
  geom_line()+
  annotate("text",x=db3$year[103], y=max(db3$flt01)+30, label= "2010",size=4)+
  #geom_smooth(method = "lm") +
  geom_rect(data=db3,aes(xmin = year[97], xmax = year[109], ymin = -Inf, ymax = Inf),
            fill = "lightgray", alpha = 0.02,linetype=0)+
  xlab("Time (years)") + 
  ylab("Inc. rate per 100000 inhab.")+
  guides(col=guide_legend(title="Vaccine"))+
  scale_x_continuous(breaks = seq(2003,2019,2))+
  scale_color_manual(breaks = c(0, 1),
  labels = lab,values=col)+
    theme(plot.title = element_text(size = 17),
          legend.text = element_text(size = 17),
          legend.title = element_text(size = 17),
          axis.title = element_text(size = 19),
          axis.text = element_text(size = 18))+
  ggtitle(label = "1 year old age group")



b<-ggplot(data=db3,aes(x=year,y=f0104,colour=as.factor(vac)))+
  geom_line()+
  annotate("text",x=db3$year[103], y=max(db3$f0104)+30, label= "2010",size=4)+
  #geom_smooth(method = "lm") +
  geom_rect(data=db3,aes(xmin = year[97], xmax = year[109], ymin = -Inf, ymax = Inf),
            fill = "lightgray", alpha = 0.02,linetype=0)+
            xlab("Time (years)") + 
            ylab("Inc. rate per 100000 inhab.")+
            guides(col=guide_legend(title="Vaccine"))+
            scale_x_continuous(breaks = seq(2003,2019,2))+
  scale_color_manual(breaks = c(0, 1),
  labels = lab,values=col)+
    theme(plot.title = element_text(size = 17),
          legend.text = element_text(size = 17),
          legend.title = element_text(size = 17),
          axis.title = element_text(size = 19),
          axis.text = element_text(size = 18))+
  ggtitle(label = "1-4 years old age group")

  
  
  
c<-ggplot(data=db3,aes(x=year,y=f0509,colour=as.factor(vac)))+
  geom_line()+
  annotate("text",x=db3$year[103], y=max(db3$f0509)+5, label= "2010",size=4)+
  #geom_smooth(method = "lm") +
  geom_rect(data=db3,aes(xmin = year[97], xmax = year[109], ymin = -Inf, ymax = Inf),
            fill = "lightgray", alpha = 0.02,linetype=0)+
            guides(col=guide_legend(title="Vaccine"))+
            scale_x_continuous(breaks = seq(2003,2019,2))+
  scale_color_manual(breaks = c(0, 1),
  labels = lab,values=col)+
   labs(x = "Time (years)", y = "Inc. rate per 100000 inhab.") +
    theme(plot.title = element_text(size = 17),
          legend.text = element_text(size = 17),
          legend.title = element_text(size = 17),
          axis.title = element_text(size = 19),
          axis.text = element_text(size = 18))+
  ggtitle(label = "5-9 years old age group")

  
jpeg(here("Figures","Fig1.jpg"), height = 10, width = 20, units = 'in', res=300)

theme_set(theme_minimal())
grid.arrange(a,b,c, nrow = 2)

dev.off()


d<-ggplot(data=db3,aes(x=year,y=f1019,colour=as.factor(vac)))+
  geom_line()+
  #geom_jitter(width = 0.5)+
  annotate("text",x=db3$year[102], y=max(db3$f1019)+2, label= "2010",size=4)+
  #geom_smooth(method = "lm") +
  geom_rect(data=db3,aes(xmin = year[97], xmax = year[109], ymin = -Inf, ymax = Inf),
            fill = "lightgray", alpha = 0.02,linetype=0)+
            guides(col=guide_legend(title="Vaccine"))+
            scale_x_continuous(breaks = seq(2003,2019,2))+
  scale_color_manual(breaks = c(0, 1),
  labels = lab,values=col)+
   labs(x = "Time (years)", y = "Inc. rate per 100000 inhab.") +
    theme(plot.title = element_text(size = 17),
          legend.text = element_text(size = 17),
          legend.title = element_text(size = 17),
          axis.title = element_text(size = 19),
          axis.text = element_text(size = 18))+
  ggtitle(label = "10-19 years old age group")


  
e<-ggplot(data=db3,aes(x=year,y=f2059,colour=as.factor(vac)))+
  geom_line()+
  #geom_jitter(width = 0.5)+
  annotate("text",x=db3$year[102], y=max(db3$f2059)+2, label= "2010",size=4)+
  #geom_smooth(method = "lm") +
  geom_rect(data=db3,aes(xmin = year[97], xmax = year[109], ymin = -Inf, ymax = Inf),
            fill = "lightgray", alpha = 0.02,linetype=0)+
            guides(col=guide_legend(title="Vaccine"))+
            scale_x_continuous(breaks = seq(2003,2019,2))+
  scale_color_manual(breaks = c(0, 1),
  labels = lab,values=col)+
   labs(x = "Time (years)", y = "Inc. rate per 100000 inhab.") +
    theme(plot.title = element_text(size = 17),
          legend.text = element_text(size = 17),
          legend.title = element_text(size = 17),
          axis.title = element_text(size = 19),
          axis.text = element_text(size = 18))+
  ggtitle(label = "20-59 years old age group")

  
 

f<-ggplot(data=db3,aes(x=year,y=f6079,colour=as.factor(vac)))+
  geom_line()+
  #geom_jitter(width = 0.5)+
  annotate("text",x=db3$year[102], y=max(db3$f6079)+2, label= "2010",size=4)+
  #geom_smooth(method = "lm") +
  geom_rect(data=db3,aes(xmin = year[97], xmax = year[109], ymin = -Inf, ymax = Inf),
            fill = "lightgray", alpha = 0.02,linetype=0)+
            guides(col=guide_legend(title="Vaccine"))+
            scale_x_continuous(breaks = seq(2003,2019,2))+
  scale_color_manual(breaks = c(0, 1),
  labels = lab,values=col)+
   labs(x = "Time (years)", y = "Inc. rate per 100000 inhab.") +
    theme(plot.title = element_text(size = 17),
          legend.text = element_text(size = 17),
          legend.title = element_text(size = 17),
          axis.title = element_text(size = 19),
          axis.text = element_text(size = 18))+
    ggtitle(label = "60-79 years old age group")

  
g<-ggplot(data=db3,aes(x=year,y=fmm80,colour=as.factor(vac)))+
  geom_line()+
  #geom_jitter(width = 0.5)+
  annotate("text",x=db3$year[102], y=max(db3$fmm80)+2, label= "2010",size=4)+
  #geom_smooth(method = "lm") +
  geom_rect(data=db3,aes(xmin = year[97], xmax = year[109], ymin = -Inf, ymax = Inf),
            fill = "lightgray", alpha = 0.02,linetype=0)+
            guides(col=guide_legend(title="Vaccine"))+
            scale_x_continuous(breaks = seq(2003,2019,2))+
  scale_color_manual(breaks = c(0, 1),
  labels = lab,values=col)+
   labs(x = "Time (years)", y = "Inc. rate per 100000 inhab.") +
    theme(plot.title = element_text(size = 17),
          legend.text = element_text(size = 17),
          legend.title = element_text(size = 17),
          axis.title = element_text(size = 19),
          axis.text = element_text(size = 18))+
    ggtitle(label = ">80 years old age group")

jpeg(here("Figures","Fig2.jpg"), height = 10, width = 20, units = 'in', res=300)

theme_set(theme_minimal())
grid.arrange(d,e,f,g, nrow = 2)

dev.off()


```


## Auto-correlation
```{r, include=FALSE}

ts_flt01 <- ts(db$flt01, start=c(2002,1), end=c(2019, 12), frequency = 12 )
ts_f0104 <- ts(db$f0104, start=c(2002,1), end=c(2019, 12), frequency = 12 )
ts_f0509 <- ts(db$f0509, start=c(2002,1), end=c(2019, 12), frequency = 12 )
ts_f1019 <- ts(db$f1019, start=c(2002,1), end=c(2019, 12), frequency = 12 )

ts_f2059 <- ts(db$f2059, start=c(2002,1), end=c(2019, 12), frequency = 12 )

ts_f6069 <- ts(db$f6069, start=c(2002,1), end=c(2019, 12), frequency = 12 )
ts_f7079 <- ts(db$f7079, start=c(2002,1), end=c(2019, 12), frequency = 12 )
ts_fmm80 <- ts(db$fmm80, start=c(2002,1), end=c(2019, 12), frequency = 12 )


plot(acf((ts_flt01),lag=100))
plot(acf((ts_f0104),lag=100))
plot(acf((ts_f0509),lag=100))
plot(acf((ts_f1019),lag=100))
plot(acf((ts_f2059),lag=100))
plot(acf((ts_f6069),lag=100))
plot(acf((ts_f7079),lag=100))
plot(acf((ts_fmm80),lag=100))


plot(pacf((ts_flt01)),lag=100)
plot(pacf((ts_f0104)))
plot(pacf((ts_f0509)))
plot(pacf((ts_f1019)))

plot(pacf((ts_f2059)))
plot(pacf((ts_f6069)))
plot(pacf((ts_f7079)))
plot(pacf((ts_fmm80)))

```

## Box-plots of sazonality

```{r ,include=FALSE}

boxplot(ts_flt01~cycle(ts_flt01),ylab='Incid?ncia',xlab='Meses')
boxplot(ts_f0104~cycle(ts_f0104),ylab='Incid?ncia',xlab='Meses')
boxplot(ts_f0509~cycle(ts_f0509),ylab='Incid?ncia',xlab='Meses')
boxplot(ts_f1019~cycle(ts_f1019),ylab='Incid?ncia',xlab='Meses')
boxplot(ts_f2059~cycle(ts_f2059),ylab='Incid?ncia',xlab='Meses')
boxplot(ts_f6069~cycle(ts_f6069),ylab='Incid?ncia',xlab='Meses')
boxplot(ts_f7079~cycle(ts_f7079),ylab='Incid?ncia',xlab='Meses')
boxplot(ts_fmm80~cycle(ts_fmm80),ylab='Incid?ncia',xlab='Meses')
```



## "Augmented Dickey-Fuller" test 
**a** $H_0:$ reject the null hypotheses (*i.e.* no stationary), concluding stochasticity in the series.

```{r, include=FALSE}
adf.test(ts_flt01)$p.value
adf.test(ts_f0104)$p.value
adf.test(ts_f0509)$p.value
adf.test(ts_f1019)$p.value
adf.test(ts_f2059)$p.value
adf.test(ts_f6069)$p.value
adf.test(ts_f7079)$p.value
adf.test(ts_fmm80)$p.value
```


## Tests for cicles: G statistics of Fisher
**a*** $H_0:$ states that there is no seasonality, thus, we reject the null hypothesis and conclude that there is seazonality.
```{r, include=FALSE}

fisher.g.test (ts_flt01)
fisher.g.test (ts_f0104)
fisher.g.test (ts_f0509)
fisher.g.test (ts_f1019)

fisher.g.test (ts_f2059)
fisher.g.test (ts_f6069)
fisher.g.test (ts_f7079)
fisher.g.test (ts_fmm80)
```


# Dinamic regressions
**a** see R support about the library dynlm (http://cran.r-project.org/web/packages/dynlm/dynlm.pdf)


## Multivariable with trend, seazonality and vaccin
```{r, echo=FALSE, fig.width=13}

vac<-c(rep(0,108),rep(1,108))
vac<-ts(vac,start=c(2002,1), end=c(2019, 12), frequency = 12) 


time_set<-cbind(db[,c(1,2,3,4,14,15,16,13)],vac)

time_set2<-cbind(db2[,c(1,2,3,4,14,15,16,13)],vac)



strCols = names(time_set)
formula4 <- list(); model4 <- list()
for (j in 2:8){

  
  formula4[[j]] = paste0(strCols[j], " ~ ", 'Full')
  model4[[j]] =( dynlm ( 
         ts(time_set[,j],start=c(2002,1), end=c(2019, 12), frequency = 12)
          ~ 
           season(ts(time_set[,j],start=c(2002,1), end=c(2019, 12), frequency = 12))+
           trend(ts(time_set[,j],start=c(2002,1), end=c(2019, 12), frequency = 12))*factor(vac)+
           ts(time_set2[,j],start=c(2002,1), end=c(2019, 12), frequency = 12)+
           L(ts(time_set[,j],start=c(2002,1), end=c(2019, 12), frequency = 12) ,1:2)+
           L(ts(time_set2[,j],start=c(2002,1), end=c(2019, 12), frequency = 12),1)
         ))

}



resumo<-list()
for (j in 2:8){

  
    resumo[[j]]=summary(model4[[j]]) 
}

names(resumo)<-formula4

for(j in 2:8){
rownames(resumo[[j]]$coefficients)<-c("Intercepto","Fev","Mar","Abr","Mai","Jun","Jul","Ago","Set","Out","Nov","Dez","Trend","vac","Confoun","L1","L2","Lc1","vac*trend")
}



jun=list(resumo[[2]]$coefficients,resumo[[3]]$coefficients,resumo[[4]]$coefficients,resumo[[5]]$coefficients,resumo[[6]]$coefficients,resumo[[7]]$coefficients,resumo[[8]]$coefficients)

names(jun)<-c(names(resumo)[[2]],names(resumo)[[3]],names(resumo)[[4]],names(resumo)[[5]],names(resumo)[[6]],names(resumo)[[7]],names(resumo)[[8]])


write.csv(jun,here("Output", "incidence.csv"))

#jun1<-as.data.frame(jun)


kable(jun,digits = 2, "html") %>%
  kable_styling(full_width = F)


descrip<-cbind(
round(c(jun$`flt01 ~ Full`[c(13),c(1)],jun$`flt01 ~ Full`[c(19),c(1)],jun$`flt01 ~ Full`[c(19),c(4)]),3),
round(c(jun$`f0104 ~ Full`[c(13),c(1)],jun$`f0104 ~ Full`[c(19),c(1)],jun$`f0104 ~ Full`[c(19),c(4)]),3),
round(c(jun$`f0509 ~ Full`[c(13),c(1)],jun$`f0509 ~ Full`[c(19),c(1)],jun$`f0509 ~ Full`[c(19),c(4)]),3),

round(c(jun$`f1019 ~ Full`[c(13),c(1)],jun$`f1019 ~ Full`[c(19),c(1)],jun$`f1019 ~ Full`[c(19),c(4)]),3),
round(c(jun$`f2059 ~ Full`[c(13),c(1)],jun$`f2059 ~ Full`[c(19),c(1)],jun$`f2059 ~ Full`[c(19),c(4)]),3),
round(c(jun$`f6079 ~ Full`[c(13),c(1)],jun$`f6079 ~ Full`[c(19),c(1)],jun$`f6079 ~ Full`[c(19),c(4)]),3),

round(c(jun$`fmm80 ~ Full`[c(13),c(1)],jun$`fmm80 ~ Full`[c(19),c(1)],jun$`fmm80 ~ Full`[c(19),c(4)]),3)
)


```



## Residuals and auto-correlation
```{r ,include=FALSE}
plot(model4[[2]])
plot(model4[[3]])
plot(model4[[4]])
plot(model4[[5]])
plot(model4[[6]])
plot(model4[[7]])
plot(model4[[8]])


plot(acf(model4[[2]]$residuals),lag=100)
plot(acf(model4[[3]]$residuals),lag=100)
plot(acf(model4[[4]]$residuals),lag=100)
plot(acf(model4[[5]]$residuals),lag=100)
plot(acf(model4[[6]]$residuals),lag=100)
plot(acf(model4[[7]]$residuals),lag=100)
plot(acf(model4[[8]]$residuals),lag=100)


plot(pacf(model4[[2]]$residuals))
plot(pacf(model4[[3]]$residuals))
plot(pacf(model4[[4]]$residuals))
plot(pacf(model4[[5]]$residuals))
plot(pacf(model4[[6]]$residuals))
plot(pacf(model4[[7]]$residuals))
plot(pacf(model4[[8]]$residuals))


vif(model4[[2]])
vif(model4[[3]])
vif(model4[[4]])
vif(model4[[5]])
vif(model4[[6]])
vif(model4[[7]])
vif(model4[[8]])

dwtest(model4[[2]])
dwtest(model4[[3]])
dwtest(model4[[4]])
dwtest(model4[[5]])
dwtest(model4[[6]])
dwtest(model4[[7]])
dwtest(model4[[8]])

```


